%Below we provide instructions about how to install \goal{} (if not yet installed), and our \goal-plugin containing the implementation of the Fribourg construction.

Below we provide instructions for obtaining and installing our plugin. We also explain how to install and use the \goal{} tool, in case it is not yet installed.

\subsubsection{\goal}
\goal{} can be downloaded from \url{http://goal.im.ntu.edu.tw/wiki/doku.php} (requires registration of name and e-mail address). After the download, \goal{} does not require an actual installation. It is run directly from within the donwloaded directory tree. The root directory \texttt{<root>} of this directory tree has the format \texttt{GOAL-YYYYMMDD}, where \texttt{YYYYMMDD} is the release date of the downloaded \goal{} version.

There are two executables for running \goal:

\begin{itemize}
\item \texttt{<root>/goal}
\item \texttt{<root>/gc}
\end{itemize}

The first one, \texttt{<root>/goal}, starts the graphical user interface of \goal. The second one, \texttt{<root>/gc} is the entry point to the command line interface of \goal. For example, running \texttt{<root>/gc help complement} shows the help page for \goal's complementation command.

\textbf{Note for Mac OS X users:} the \goal{} website provides a special Mac OS X version that can be downloaded as a \texttt{.dmg} file. This file installs \goal{} as an application bundle under \texttt{/Applications/GOAL.app}. The root directory \texttt{<root>} of the \goal{} directory tree is in this case \texttt{GOAL.app/Contents/Resources/Java}.


\subsubsection{Fribourg Construction Plugin}

For obtaining and installing our plugin to an existing \goal{} installation, execute the following steps:

\begin{enumerate}
\item Download the plugin from \url{https://frico.s3.amazonaws.com/goal_plugins/ch.unifr.goal.complement.zip}
\item Unzip the donwloaded file \texttt{ch.unifr.goal.complement.zip}
\item Move the unzipped directory \texttt{ch.unifr.goal.complement} to \texttt{<root>/plugins}
\end{enumerate}

This is all. When \goal{} is started for the next time, the Fribourg construction will be included in the graphical user interface, as well as in the command line interface of \goal.

\textbf{Important note:} the Fribourg construction plugin requires \goal{} version 2014--11--17 or later. The plugin does not work with older versions of \goal. This is because between the 2014--08--08 and the 2014--11--17 version, the plugin interfaces of \goal{} have changed.

Note that our additional plugin \texttt{ch.unifr.goal.util}, which contains two commands for testing completeness and universality of automata (only for the command line interface), can be downloaded from \url{https://frico.s3.amazonaws.com/goal_plugins/ch.unifr.goal.util.zip} and installed according to the above instructions.
